Perihelion advance of Mercury

### 2020-10-07 perihelion_advance_of_mercury.Rmd
identify perihelion instead of aphelion

In the above plot, note the differing axis scales.  The start point is
shown as a solid circle, subseqent points shown with open circles.
Lines connect consecutive points in the simulation.  It is difficult
to see evidence for precession in this plot because the points
correspond to regular snapshots of Mercury's position in successive
orbits, and the time evolution is not shown.  To proceed we need to
find the aphelion (or an approximation to it) somehow, and show that
the aphelion systematically increases with time.  Given the numerics,
we find points on the simulated trajectory which are a local maximum
distance from the sun.  This is straightfoward numerically, we
identify iterations $i$ (which have radius `r[i]`) satisfying both
`r[i-1]<r[i]` and `r[i]>r[i+1]`:

```{r,label=orbitalangleatperihelion}
jj1 <- c(NA,out$rdot)
jj2 <- c(out$rdot,NA)
change <- (jj1>0) & (jj2<0)

@@

plot(tee,phi-pi,main='orbital angle of Mercury at perihelion',xlab='time (seconds)',ylab='angle of perihelion')
```

>>>>

In the above plot, note the differing axis scales.  The start point is
shown as a solid circle, subseqent points shown with open circles.
Lines connect consecutive points in the simulation.  It is difficult
to see evidence for precession in this plot because the points
correspond to regular snapshots of Mercury's position in successive
orbits, and the time evolution is not shown.

To proceed we need to find the perihelion (or an approximation to it)
somehow, and show that the estimated perihelion systematically
increases with time.  Given the numerics, we find points on the
simulated trajectory which are a local minimum distance from the sun.
This is straightfoward numerically; we identify iterations $i$ (which
have radius `r[i]`) satisfying both `r[i-1]<r[i]` and `r[i]>r[i+1]`:

```{r,label=orbitalangleatperihelion}
jj1 <- c(NA,out$rdot)
jj2 <- c(out$rdot,NA)
change <- (jj1<0) & (jj2>0)

@@

plot(tee,phi,main='orbital angle of Mercury at perihelion',xlab='time (seconds)',ylab='angle of perihelion')
```
